### 2022-11-30 HardwareX - Spectrophotometer/sec4_bom.tex

\section{Bill of materials}

% For a complex Bill of Materials, the complete Bill of Materials (editable spreadsheet file e.g., ODS file type or PDF file) can be uploaded in an open access online location such as the Open Science Framework repository. Include the link here. Alternatively, the Bill of Materials can be uploaded at the time of submission on the online Elsevier submission interface as supplementary material.

% > To make it easy to tell which item in the Bill of Materials corresponds to which component in your design file(s), use matching designators in both places, or otherwise explain the correspondence.

% > For material type, select from: Metal, semi-conductor, ceramic, polymer, biomaterial, organic, inorganic, composite, nanomaterial, semiconductor, non-specific, or other  
\textit{For a complex Bill of Materials, the complete Bill of Materials (editable spreadsheet file e.g., ODS file type or PDF file) can be uploaded in an open access online location such as the Open Science Framework repository. Include the link here. Alternatively, the Bill of Materials can be uploaded at the time of submission on the online Elsevier submission interface as supplementary material.}
\begin{itemize}
\item \textit{To make it easy to tell which item in the Bill of Materials corresponds to which component in your design file(s), use matching designators in both places, or otherwise explain the correspondence.}
\item \textit{For material type, select from: Metal, semi-conductor, ceramic, polymer, biomaterial, organic, inorganic, composite, nanomaterial, semiconductor, non-specific, or other} 
\end{itemize}

%\begin{landscape}
\tabulinesep=1ex
\begin{tabu} to \linewidth {|l|l|l|l|l|l|l|}
\hline
\textbf{Designator} & \textbf{Component} & \textbf{Number} & \textbf{Cost per unit - USD} & \textbf{Total cost - USD} & \textbf{Source of materials} & \textbf{Material type} \\\hline

%Insert items here
\textit{Designator 1} & \textit{Name of Component 1} & \textit{Number of units} & \textit{Cost per unit} & \textit{Total cost} & \textit{Source} & \textit{Material type} \\\hline
\textit{Designator 2} & \dots & \dots & \dots & \dots & \dots & \dots \\\hline
% Designator 3 & Name of Component 2 & Number of units & Cost per unit & Total cost & Source of materials & Material type \\\hline
\end{tabu}
%\end{landscape}