### 2022-11-30 HardwareX - Spectrophotometer/sec5_buildinstr.tex

\section{Build instructions}

%Provide detailed, step by step instructions for the construction of the reported hardware include all necessary information for reproducing the submitted hardware.
% > Explain and, when possible, characterize design decisions. Including design alternatives if they exist. 
% > Use visual instructions such as schematics, images, and videos. 
% > Clearly reference design files and component parts described in the Design File Summary and Bill of Materials. 
% >Highlight potential safety concerns that may arise

\textit{Provide detailed, step by step instructions for the construction of the reported hardware
 include all necessary information for reproducing the submitted hardware.
\begin{itemize}
\item Explain and, when possible, characterize design decisions. Including design alternatives if they exist. 
\item Use visual instructions such as schematics, images, and videos. 
\item Clearly reference design files and component parts described in the Design File Summary and Bill of Materials. 
\item Highlight potential safety concerns that may arise
\end{itemize}}