### 2022-11-30 HardwareX - Spectrophotometer/sec6_operation.tex

\section{Operation instructions}

%Provide detailed instructions for the safe and proper operation of the hardware. 
%> Step-by-step operational instructions for operating the hardware. 
%> Use visual instructions as necessary. 
%> Highlight potential safety hazards.

\textit{Provide detailed instructions for the safe and proper operation of the hardware. 
\begin{itemize}
\item Step-by-step operational instructions for operating the hardware. 
\item Use visual instructions as necessary. 
\item Highlight potential safety hazards.
\end{itemize}}