### 2022-11-30 HardwareX - Spectrophotometer/sec7_validation.tex

\section{Validation and characterization}

%Demonstrate the operation of the hardware and characterize its performance over relevant critical metrics
%> Demonstrate the use of the hardware for a relevant use case. 
%> If possible, characterize performance of the hardware over operational parameters. 
%> Create a bulleted list that describes the capabilities (and limitations) of the hardware. For example consider descriptions of load, operation time, spin speed, coefficient of variation, accuracy, precision and etc.

\textit{Demonstrate the operation of the hardware and characterize its performance over relevant critical metrics
\begin{itemize}
\item Demonstrate the use of the hardware for a relevant use case. 
\item If possible, characterize performance of the hardware over operational parameters. 
\item Create a bulleted list that describes the capabilities (and limitations) of the hardware. For example consider descriptions of load, operation time, spin speed, coefficient of variation, accuracy, precision and etc. 
\end{itemize}}