### 2022-11-30 HardwareX - Spectrophotometer/sec8_acknow.tex

\newpage
\section{Acknowledgements}
This research works was funded by Lembaga Pengembangan Inovasi dan Kewirausahaan, Institut Teknologi Bandung (LPIK ITB) under the grant "Program Penguatan Inovasi 2021" number 289L/IT1.A/SK-KP/2021 and School of Electrical Engineering and Informatics, Institut Teknologi Bandung, under the grant "Program Penelitian, Pengabdian kepada Masyarakat dan Inovasi, Institut Teknologi Bandung, 2021 (P2MI-ITB 2021)" number 917/IT1.C12/KP/2021.

The article processing charge (APC) was funded by School of Electrical Engineering, Telkom University.